Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambio en nombre del doc
cambios en la intro

%% Group authors per affiliation:
\author{Mauricio Gonzalez Soto}
\author{Luis E. Sucar Succar}
\author{Hugo J. Escalante Balderas}
\address{Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica, Luis Enrique Erro 1, Santa Maria Tonanzintla, Mexico}

@@

%\tableofcontents
\section{Introduction}
\label{intro}
Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{tversky1977causal,waldmann2013causal,danks2014unifying,neil2019causality}. It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}. 
Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.

Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent, located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward. Optimal policies, which are prescriptions of which action must be taken at some state $s$, satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem. Such solution is equivalent to maximizing \textit{expected utility} \citep{webb2007game}, which is the decision criterion prescribed by the most commonly used decision-making theories: von Neumann-Morgenstern's and Savage's. Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations. In fact, it can be argued that RL is not a causal problem in itself \cite{gonzalezsoto2019reinforcement}.

>>>>

%% Group authors per affiliation:
\author[1]{Mauricio Gonzalez-Soto}
\ead{mauricio@inaoep.mx}

\author[1]{L. Enrique Sucar}
\ead{esucar@inaoep.mx}

\author[1]{Hugo J. Escalante}
\ead{hugojair@inaoep.mx}

\address[1]{Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica ,Luis Enrique Erro 1, Santa Maria Tonanzintla, M\'exico.}

@@

%\tableofcontents
\section{Introduction}
\label{intro}
Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{tversky1977causal,waldmann2013causal,danks2014unifying,neil2019causality}. It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}. 
Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal,lagnado2007beyond,hagmayer2008causal,hagmayer2009decision,hagmayer2017causality}.

Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward. Optimal policies are prescriptions of which action must be taken at some state $s$ that satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem. Such solution is equivalent to maximizing the \textit{expected utility} \citep{shoham2008multiagent,webb2007game}. Maximization of expected utility is the criterion prescribed by the most common used decision-making theories: von Neumann-Morgenstern's and Savage's. Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations. In fact, it can be argued that RL is not a causal problem in itself \cite{gonzalezsoto2019reinforcement}.
